Tomographic galaxy clustering with the Subaru Hyper Suprime-Cam first year public data release

### 2019-10-20 doc/Paper/main.tex
Additional small changes.

The past two decades have seen a revolution in our understanding of the universe and its constituents. While observations of the cosmic microwave background play a pivotal role \cite{Planck:2018}, they are only indirectly sensitive to physics of the low redshift universe, in particular to the expansion history and growth, which are uniquely sensitive to properties of dark energy and modified gravity \cite{1903.12016}.

@@

Together with transformational sensitivity increases in the Stage IV surveys, the challenges of understanding and controlling systematic errors are becoming considerably more difficult \cite{1808.07335}. For photometric surveys, the deep surveys sensitive to many more sources inevitably come with crowded fields where effects of blending affect a significant fraction of all sources. The increased blending leads to new uncertainties in isolating sources, measuring their fluxes (required for photometric redshifts) and inferring their shear estimates. These effects in turn lead to subtle sample selection effects, which are amplified by the interaction of the point spread function with blending \cite{1708.01533,1905.01324,1907.10572}. But even with perfect measurements of fluxes, our ability to infer source redshift distributions is hampered by incompleteness in spectroscopic samples used to calibrate redshifts \cite{1903.09325}. Finally, at the depths of Stage IV photometric surveys, instrumental and observational effects, such as scattered light from bright objects, errors in star-galaxy separtation, reddening by milky way dust and others will imprint spurious fluctuations on the observed galaxy density field. These fluctuations will, if not correctly accounted for, masquerade as intrinsic large-scale fluctuations in the cosmic density fields. Therefore, Stage IV photometric experiments are likely to be limited by our ability to control systematics rather than the intrinsic statistical power of the survey.

@@

In this work, we present an analysis of photometric galaxy clustering using HSC DR1 data. We split the data into four tomographic redshift bins and measure both auto- and cross power spectra, correcting for observational systematics using mode deprojection. We then use the halo model coupled with a halo occupation distribution to fit the data and derive constraints on astrophysical parameters, marginalizing over photometric redshift uncertainties. To ensure the stability of our results, we perform a suite of complementary analyses and find our constraints to be robust to photometric redshift uncertainties. We additionally perform extended analyses in which we compute constraints on magnification in our sample and allow for variations in the $\Lambda$CDM cosmological parameters respectively. Finally, we use our data to derive a simple fitting function for the linear galaxy bias as a function of redshift and limiting magnitude of the sample. In our analysis, we make several simplifying approximations; in particular we employ a simple semi-analytical halo model to model the data, which will most certainly not be sophisticated enough for the full sky LSST analysis. However, we find that it suffices for the statistical precision of data at hand, but we view the use of halo model as a ``good enough'' substitute for the final modeling, which will likely include bias expansion approaches \cite{0902.0991,1402.5916,1611.09787} and power spectrum emulators \cite{1804.05865,1705.03388}.

@@

  We use data from the Hyper Suprime-Cam Subaru Strategic Program first data release (HSC DR1 hereon)\footnote{https://hsc-release.mtk.nao.ac.jp}. The release is extensively documented in \cite{2018PASJ...70S...8A}, and we only provide the details of the galaxy sample and associated data used for our clustering analysis.

  HSC-SSP is a photometric galaxy survey that has been awarded 300 nights on the Subaru Telescope starting in 2014. DR1 includes data from 61.5 nights covering and comes with fields of three depths: Wide (108 square degrees to $i\sim 26.4$), Deep (26 square degrees to $i\sim26.5$) and UltraDeep (4 square degrees to $i\sim 27$). In this work we focus on the Wide field, which has been observed in five broadband filters ($grizy$) and is distributed among 7 fields between 5 and 20 square degrees each. The image quality is impressive at median seeing of around 0.6 arcsec, which is considerably better than other comparable surveys (e.g. DES with around 0.9 arc sec) and also likely better than median seeing that will be achieved by LSST. The data are processed with \texttt{hscPipe} \cite{2018PASJ...70S...5B} and  available to the community as a publicly available database.

  We use a magnitude-limited sample constructed from the WIDE DR1 sample by imposing data cuts that are similar to those used to create the HSC shear catalog \cite{2018PASJ...70S..25M}. The exact cuts are shown in Table~\ref{tab:cuts_summary}, and can be summarized as follows: besides a minimal set of quality cuts (selecting only primary detections with well-measured fluxes in all bands, removing objects near bad pixels, deblender artifacts, etc.), we impose an overall apparent magnitude cut in the extinction-corrected band $i_{\rm corr}<24.5$. This choice was based on a study of the survey depth-completeness relation in order to select a homogeneous and complete sample of high-confidence ($>10\sigma$) detections (see Section \ref{ssec:methods.syst}). We also select only objects with significant detections ($>5\sigma$) in at least two of the 4 remaining bands ($g,\,r,\,z,\,y$), and remove all objects classified as stars by the data reduction pipeline, using the "extendedness`` classifier as described in \cite{2018PASJ...70S..25M,2018PASJ...70S...5B}. The resulting sample consists of 9,734,457 objects and covers $\sim88$ square degrees distributed across the 6 HSC DR1 fields, as described in Table \ref{tab:field_summary}.

@@

Our basic method is to measure the two point function of all possible combinations of galaxy density fluctuation fields. We work in the Fourier domain, so our basic quantity is the angular power spectrum $C_\ell$. Given four tomographic samples, our measurement consists of four auto power spectra and six cross-power spectra. In this section we discuss how we construct maps of galaxy density fluctuations and associate maps of potential systematics , how we use these to measure the power spectra and its covariance matrix and finally how these are modeled within the context of halo model.

@@

    To generate maps of all the quantities described in this section we make use of a rectangular pixelization scheme using the Plate Carr\'ee projection (labelled {\tt CAR} in the World Coordinate System standard \cite{2002A&A...395.1077C}). In this case, pixels are simply defined by equal intervals of colatitude $\theta$ and azimuth $\phi$. To minimize the distortions caused by the flat-sky approximation, we place the projection reference point (i.e. a point in the equator $\theta=\pi/2$) at the center of each field. We use square pixels of size $\alpha_{\rm pix}=0.6$ arcmin on a side, corresponding to a Nyquist frequency $\ell_{\rm Nyquist}=18,000$. The maps are defined on a rectangular patch large enough to cover all objects in each field, leaving a buffer of 10 masked pixels on all edges to avoid boundary effects when computing the FFTs.

@@

    Following \cite{2011ApJ...736...59Z}, we parametrise the number of centrals and satellites as a function of mass as:
    \begin{align}
      &\bar{N}_c(M)=\frac{1}{2}\left[1+{\rm erf}\left(\frac{\log(M/M_{\rm min})}{\sigma_{{\rm ln}M}}\right)\right],\\
      &\bar{N}_s(M)=\Theta(M-M_0)\left(\frac{M-M_0}{M_1'}\right)^\alpha,
    \end{align}
    where $\Theta(x)$ is the Heavyside step function. Furthermore, we assume that $N_c$ follows a Bernoulli distribution with probability $p=\bar{N}_c$, and that the number of satellites is Poisson-distributed with mean $\bar{N}_s$. Finally, we model the distribution of satellites to follow that of the dark matter, and therefore $u_s$ is a Navarro-Frenk-White profile, given by:
    \begin{equation}
      u_s(k|M)=\frac{\sin x\left[{\rm Si}\left((1+c)\,x\right)-{\rm Si}(x)\right]+\cos x\left[{\rm Ci}\left((1+c)x\right)-{\rm Ci}(x)\right]-\frac{\sin(cx)}{(1+c)x}}{{\rm ln}(1+c)-\frac{c}{1+c}},
    \end{equation}
    where $x=\sfrac{k R_\Delta}{c}$, $R_\Delta$ is the halo radius, $c=c(M)$ is the concentration parameter, and ${\rm Si}/{\rm Ci}$ are the sine and cosine integral functions. We define $R_\Delta$ as the radius that encloses $\Delta=200$ times the background matter density. The concentration-mass relation $c(M)$ also depends on the choice of $\Delta$, and we follow the parametrization of \cite{Duffy:2008}. In this work, we model the halo mass function $\sfrac{\mathrm{d}n}{\mathrm{d}M}$ following \cite{Tinker:2010}. Furthermore, we employ the concentration-mass relation $c(M)$ (which depends on the choice of $\Delta$) derived by \cite{Duffy:2008}.

@@

  \subsection{Parameter constraints}\label{ssec:methods.constr}
    In order to derive constraints on HOD, cosmological and systematics parameters, we assume the joint likelihood of all auto- and cross power spectra to be Gaussian 
    \begin{align}
      \mathscr{L}(D \vert \theta) = \frac{1}{[(2\pi)^{d}\det{\mathsf{C}}]^{\sfrac{1}{2}}} e^{-\frac{1}{2}(\mathbf{C}^{\mathrm{obs}}_{\ell}-\mathbf{C}^{\mathrm{theor}}_{\ell})^{\mathrm{T}}\mathsf{C}^{-1}(\mathbf{C}^{\mathrm{obs}}_{\ell}-\mathbf{C}^{\mathrm{theor}}_{\ell})}.
      \label{eq:likelihood}
    \end{align}
    The quantity $\mathsf{C}$ denotes the joint covariance matrix, which we estimate analytically as described in Sec.~\ref{sssec:methods.theory.covar} and is kept constant during parameter estimation \citep{2019OJAp....2E...3K}.
    We sample the likelihood in a Monte Carlo Markov Chain (MCMC) using the publicly available code \texttt{CosmoHammer} \cite{Akeret:2013}, which is based on \texttt{emcee} \cite{Foreman-Mackey2013}\footnote{We note that we have compared the MCMC results obtained using \texttt{CosmoHammer} to those obtained using a Metropolis-Hastings algorithm as implemented in \texttt{april} (https://github.com/slosar/april), finding consistent results for our test case.}. In our fiducial analysis we sample the parameter set $\boldsymbol{\theta} = \{M_{\mathrm{min}}, \allowbreak \, M_{\mathrm{min}, p}, \allowbreak \, M_{0}, \allowbreak \, M_{0, p}, \allowbreak \, M_{1}, \allowbreak \, M_{1, p}, \allowbreak \, \Delta z_{i}, \allowbreak \, z_{w, i}\}$, $i = 0, \dots ,3$, where the first six parameters describe the HOD of galaxies as outlined in Sec.~\ref{sssec:methods.theory.hod}. The remaining parameters account for photometric redshift uncertainties as described in Sec.~\ref{sssec:methods.theory.photoz_syst}. We perform additional analyses in which we separately allow for variations in the amplitude $A_{\mu}$ of the magnification bias kernel $W_{\mu}$ and the cosmological parameters $\Omega_{c}$ and $\sigma_{8}$, where $\Omega_{c}$ is the the fractional cold matter density today and $\sigma_{8}$ denotes the r.m.s. of linear matter fluctuations in spheres of comoving radius 8 $h^{-1}$ Mpc. For all sampled parameters, we assume flat, uniform priors. The sampled parameters are shown alongside their priors in Tab.~\ref{tab:params}. Unless stated otherwise, we fix all cosmological parameters to the best-fit values derived by the Planck Collaboration in 2018 using temperature, polarization and CMB lensing data, i.e. $\Omega_{b}=0.0493$, $\Omega_{c}=0.264$, $h=0.6736$, $n_{s}=0.9649$ and $\sigma_{8}=0.8111$ (see the fourth column in Tab.~2 in \cite{Planck:2018}).

>>>>

The past two decades have seen a revolution in our understanding of the universe and its constituents. While observations of the cosmic microwave background play a pivotal role \cite{Planck:2018}, they only indirectly probe the physics of the low redshift universe, in particular the expansion history and growth, which are uniquely sensitive to properties of dark energy and modified gravity \cite{1903.12016}.

@@

Together with transformational sensitivity increases in the Stage IV surveys, the challenges of understanding and controlling systematic errors are becoming considerably more difficult \cite{1808.07335}. For photometric surveys, the deep surveys sensitive to many more sources inevitably come with crowded fields where effects of blending affect a significant fraction of all sources. The increased blending leads to new uncertainties in isolating sources, measuring their fluxes (required for photometric redshifts) and inferring their shear estimates. These effects in turn lead to subtle sample selection effects, which are amplified by the interaction of the point spread function with blending \cite{1708.01533,1905.01324,1907.10572}. But even with perfect measurements of fluxes, our ability to infer source redshift distributions is hampered by incompleteness in spectroscopic samples used to calibrate redshifts \cite{1903.09325}. Finally, at the depths of Stage IV photometric surveys, instrumental and observational effects, such as scattered light from bright objects, errors in star-galaxy separation, reddening by milky way dust and others will imprint spurious fluctuations on the observed galaxy density field. These fluctuations will, if not correctly accounted for, masquerade as intrinsic large-scale fluctuations in the cosmic density fields. Therefore, Stage IV photometric experiments are likely to be limited by our ability to control systematics rather than the intrinsic statistical power of the survey.

@@

In this work, we present an analysis of photometric galaxy clustering using HSC DR1 data. We split the data into four tomographic redshift bins and measure both auto- and cross power spectra, correcting for observational systematics using mode deprojection. We then use the halo model coupled with a halo occupation distribution to fit the data and derive constraints on astrophysical parameters, marginalizing over photometric redshift uncertainties. To ensure the stability of our results, we perform a suite of complementary analyses and find our constraints to be robust to photometric redshift uncertainties. We additionally perform extended analyses in which we compute constraints on magnification in our sample and allow for variations in the $\Lambda$CDM cosmological parameters, respectively. Finally, we use our data to derive a simple fitting function for the linear galaxy bias as a function of redshift and limiting magnitude of the sample. In our analysis, we make several simplifying approximations; in particular we employ a simple semi-analytical halo model to model the data, which will most certainly not be sophisticated enough for the full sky LSST analysis. However, we find that it suffices for the statistical precision of the data at hand, but we view the use of the halo model as a ``good enough'' substitute for the final modeling, which will likely include bias expansion approaches \cite{0902.0991,1402.5916,1611.09787} and power spectrum emulators \cite{1804.05865,1705.03388}.

@@

In this work, we use data from the first data release of the Hyper Suprime-Cam Subaru Strategic Program (HSC DR1 hereon)\footnote{\texttt{https://hsc-release.mtk.nao.ac.jp}}. The release is extensively documented in \cite{2018PASJ...70S...8A}, and we only provide the details of the galaxy sample and associated data used for our clustering analysis.

  HSC-SSP is a photometric galaxy survey that has been awarded 300 nights on the Subaru Telescope starting in 2014. DR1 includes data from 61.5 nights observed to three different depths: Wide (108 square degrees to $i\sim 26.4$), Deep (26 square degrees to $i\sim26.5$) and UltraDeep (4 square degrees to $i\sim 27$). In this work we focus on the Wide field, which has been observed in five broadband filters ($grizy$) and is distributed among 7 fields  of areas varying between 5 and 20 square degrees. The image quality is impressive at median seeing of around 0.6 arcsec, which is considerably better than other comparable surveys (e.g. DES with around 0.9 arc sec) and also likely better than the median seeing that will be achieved by LSST. The data are processed with \texttt{hscPipe} \cite{2018PASJ...70S...5B} and are available to the community through a publicly available database.

  We use a magnitude-limited sample constructed from the WIDE DR1 sample by imposing data cuts that are similar to those used to create the HSC shear catalog \cite{2018PASJ...70S..25M}. The exact cuts are shown in Table~\ref{tab:cuts_summary}, and can be summarized as follows: besides a minimal set of quality cuts (selecting only primary detections with well-measured fluxes in all bands, removing objects near bad pixels, deblender artifacts, etc.), we impose an overall apparent magnitude cut in the extinction-corrected band $i_{\rm corr}<24.5$. This choice was based on a study of the survey depth-completeness relation in order to select a homogeneous and complete sample of high-confidence ($>10\sigma$) detections (see Section \ref{ssec:methods.syst}). We also select only objects with significant detections ($>5\sigma$) in at least two of the 4 remaining bands ($g,\,r,\,z,\,y$), and remove all objects classified as stars by the data reduction pipeline, using the ``extendedness'' classifier as described in \cite{2018PASJ...70S..25M,2018PASJ...70S...5B}. The resulting sample consists of 9,734,457 objects and covers $\sim88$ square degrees distributed across the 6 HSC DR1 fields, as described in Table \ref{tab:field_summary}.

@@

Our basic method is to measure the two point function of all possible combinations of galaxy density fluctuation fields. We work in the Fourier domain, so our basic quantity is the angular power spectrum $C_\ell$. Given four tomographic samples, our measurement consists of four auto power spectra and six cross-power spectra. In this section we discuss how we construct maps of galaxy density fluctuations and associated maps of potential systematics, how we use these to measure the power spectra and their covariance matrix and finally how these are modeled within the context of the halo model.

@@

    To generate maps of all the quantities described in this section, we make use of a rectangular pixelization scheme using the Plate Carr\'ee projection (labelled {\tt CAR} in the World Coordinate System standard \cite{2002A&A...395.1077C}). In this case, pixels are simply defined by equal intervals of colatitude $\theta$ and azimuth $\phi$. To minimize the distortions caused by the flat-sky approximation, we place the projection reference point (i.e. a point in the equator $\theta=\pi/2$) at the center of each field. We use square pixels of size $\alpha_{\rm pix}=0.6$ arcmin on a side, corresponding to a Nyquist frequency $\ell_{\rm Nyquist}=18,000$. The maps are defined on a rectangular patch large enough to cover all objects in each field, leaving a buffer of 10 masked pixels on all edges to avoid boundary effects when computing the FFTs.

@@

    Following \cite{2011ApJ...736...59Z}, we parametrise the number of centrals and satellites as a function of mass as:
    \begin{align}
      &\bar{N}_c(M)=\frac{1}{2}\left[1+{\rm erf}\left(\frac{\log(M/M_{\rm min})}{\sigma_{\log M}}\right)\right],\\
      &\bar{N}_s(M)=\Theta(M-M_0)\left(\frac{M-M_0}{M_1'}\right)^\alpha,
    \end{align}
    where $\Theta(x)$ is the Heavyside step function. Furthermore, we assume that $N_c$ follows a Bernoulli distribution with probability $p=\bar{N}_c$, and that the number of satellites is Poisson-distributed with mean $\bar{N}_s$. Finally, we model the distribution of satellites to follow that of the dark matter, and therefore $u_s$ is a Navarro-Frenk-White profile, given by \cite{Navarro:1996}:
    \begin{equation}
      u_s(k|M)=\frac{\sin x\left[{\rm Si}\left((1+c)\,x\right)-{\rm Si}(x)\right]+\cos x\left[{\rm Ci}\left((1+c)x\right)-{\rm Ci}(x)\right]-\frac{\sin(cx)}{(1+c)x}}{{\rm ln}(1+c)-\frac{c}{1+c}},
    \end{equation}
    where $x=\sfrac{k R_\Delta}{c}$, $R_\Delta$ is the halo radius, $c=c(M)$ is the concentration parameter, and ${\rm Si}/{\rm Ci}$ are the sine and cosine integral functions. We define $R_\Delta$ as the radius that encloses $\Delta=200$ times the background matter density. The concentration-mass relation $c(M)$ also depends on the choice of $\Delta$, and we follow the parametrization of \cite{Duffy:2008}. In this work, we model the halo mass function $\sfrac{\mathrm{d}n}{\mathrm{d}M}$ following \cite{Tinker:2010}. Furthermore, we employ the concentration-mass relation $c(M)$ (which depends on the choice of $\Delta$) derived by \cite{Duffy:2008}.

@@

  \subsection{Parameter constraints}\label{ssec:methods.constr}
    In order to derive constraints on HOD, cosmological and systematics parameters, we assume the joint likelihood of all auto- and cross power spectra to be Gaussian 
    \begin{align}
      \mathscr{L}(D \vert \theta) = \frac{1}{[(2\pi)^{d}\det{\mathsf{C}}]^{\sfrac{1}{2}}} e^{-\frac{1}{2}(\mathbf{C}^{\mathrm{obs}}_{\ell}-\mathbf{C}^{\mathrm{theor}}_{\ell})^{\mathrm{T}}\mathsf{C}^{-1}(\mathbf{C}^{\mathrm{obs}}_{\ell}-\mathbf{C}^{\mathrm{theor}}_{\ell})}.
      \label{eq:likelihood}
    \end{align}
    The quantity $\mathsf{C}$ denotes the joint covariance matrix, which we estimate analytically as described in Sec.~\ref{sssec:methods.theory.covar} and is kept constant during parameter estimation \citep{2019OJAp....2E...3K}.
    We sample the likelihood in a Monte Carlo Markov Chain (MCMC) using the publicly available code \texttt{CosmoHammer} \cite{Akeret:2013}, which is based on \texttt{emcee} \cite{Foreman-Mackey2013}\footnote{We note that we have compared the MCMC results obtained using \texttt{CosmoHammer} to those obtained using a Metropolis-Hastings algorithm as implemented in \texttt{april} (https://github.com/slosar/april), finding consistent results for our test case.}. In our fiducial analysis we sample the parameter set $\boldsymbol{\theta} = \{M_{\mathrm{min}}, \allowbreak \, M_{\mathrm{min}, p}, \allowbreak \, M_{0}, \allowbreak \, M_{0, p}, \allowbreak \, M_{1}, \allowbreak \, M_{1, p}, \allowbreak \, \Delta z_{i}, \allowbreak \, z_{w, i}\}$, $i = 0, \dots ,3$, where the first six parameters describe the HOD of galaxies as outlined in Sec.~\ref{sssec:methods.theory.hod}. The remaining parameters account for photometric redshift uncertainties as described in Sec.~\ref{sssec:methods.theory.photoz_syst}. We perform additional analyses in which we separately allow for variations in the amplitude $A_{\mu}$ of the magnification bias kernel $W_{\mu}$ and the cosmological parameters $\Omega_{c}$ and $\sigma_{8}$, where $\Omega_{c}$ is the the fractional cold matter density today and $\sigma_{8}$ denotes the r.m.s. of linear matter fluctuations in spheres of comoving radius 8 $h^{-1}$ Mpc. For all sampled parameters, we assume flat, uniform priors. The sampled parameters are shown alongside their priors in Tab.~\ref{tab:params}. The remaining HOD parameters are set to $\alpha = 1.0$ and $\sigma_{\log M} = 0.4$. Unless stated otherwise, we further fix all cosmological parameters to the best-fit values derived by the Planck Collaboration in 2018 using temperature, polarization and CMB lensing data, i.e. $\Omega_{b}=0.0493$, $\Omega_{c}=0.264$, $h=0.6736$, $n_{s}=0.9649$ and $\sigma_{8}=0.8111$ (see the fourth column in Tab.~2 in \cite{Planck:2018}).
